---
title: "PA1_template"
output: html_document
---

This my first Rmarkdown.file

**Loading and preprocessing the data**
```{r}
activity <- read.csv("C:/activity.csv", colClasses = c("numeric", "character", "numeric"))
suppressMessages(require(lattice))
activity["date"] <- as.Date(activity$date, "%Y-%m-%d")

```


**What is mean total number of steps taken per day?**

```{r}
steps_per_day <- aggregate(steps ~ date, data = activity, sum, na.rm = TRUE)
mean_steps_per_day <- mean(steps_per_day$steps)
median_steps_per_day <- median(steps_per_day$steps)

hist(steps_per_day$steps, main = "Frequency of Steps Taken Per Day",
               xlab = "Daily Step Count Range",
               ylab="Number of Days Step Count Range is Reached", col = "green3", xlim=c(0,25000), ylim=c(0,20), breaks=10)
```
*The mean steps taken per day is 10766*

```{r}
mean_steps_per_day
```
*The median steps taken per day is 10765*
```{r}
median_steps_per_day
```

**What is the average daily activity pattern?**
```{r}
time_series <- tapply(activity$steps, activity$interval %% 100 / 5 + activity$interval %/% 100 * 12 + 1, mean, na.rm = TRUE)

max_interval <- which.max(time_series)
hour_of_day_start <- (max_interval * 5) %/% 60
minute_of_hour_start <- (max_interval * 5) %% 60

hour_of_day_end <- ((max_interval + 1) * 5) %/% 60
minute_of_hour_end <- ((max_interval + 1) * 5) %% 60

am_pm_start <- "AM"
am_pm_end <- "AM"

if(hour_of_day_start > 12) {
  hour_of_day_start <- hour_of_day_start - 12
  am_pm_start <- "PM"
}

if(hour_of_day_end > 12) {
  hour_of_day_end <- hour_of_day_end - 12
  am_pm_end <- "PM"
}


plot(row.names(time_series), time_series, type = "l", xlab = "5-minute Interval (288/Day)", 
    ylab = "Daily Average (steps)", main = "Average Number of Steps During 5-minute Time Intervals In a 24-hour Day", 
    col = "blue")

max_interval

```
**Imputing missing values**
```{r}
NA_count <- sum(is.na(activity))
NA_count
suppressMessages(require(reshape))
suppressMessages(require(reshape2))
a <- dcast(activity, interval ~ date, value.var="steps", fill=0)
r <- dcast(activity, interval ~ date, value.var = "steps", fill = rowMeans(a, na.rm = TRUE))
r2 <- reshape(r, direction = "long", varying=list(names(r)[2:length(names(r))]),
                               v.names=c("steps"), timevar="date", idvar=c("interval"),
                               times=names(r)[2:length(names(r))], new.row.names=1:dim(activity)[1])

steps_per_day2 <- aggregate(steps ~ date, data = r2, sum, na.rm = TRUE)
hist(steps_per_day2$steps, main = "Frequency of Steps Taken Per Day", xlab = "Daily Step Count Range",
          ylab="Number of Days Step Count Range is Reached", col = "green3", xlim=c(0,25000), ylim=c(0,20), breaks=10)
mean_steps_per_day <- mean(steps_per_day2$steps)
median_steps_per_day <- median(steps_per_day2$steps)

mean_steps_per_day 
median_steps_per_day 
```
**re there differences in activity patterns between weekdays and weekends?**
```{r}
r2["date"] <- as.Date(r2$date, "%Y-%m-%d")
r2[(weekdays(r2$date) %in% c("Saturday", "Sunday")), "TypeOfDay"] <- "Weekend"
r2[!(weekdays(r2$date) %in% c("Saturday", "Sunday")), "TypeOfDay"] <- "Weekday"

steps <- aggregate(steps ~ interval + TypeOfDay, data = r2, mean)
names(steps) <- c("Interval", "TypeOfDay", "Steps")
steps$Interval <- steps$Interval %% 100 / 5 + steps$Interval %/% 100 * 12 + 1

splot <- xyplot(Steps ~ Interval | TypeOfDay, steps, type = "l", layout = c(1, 2),xlab = "5-minute Interval (288 per Day)", ylab = "Daily Average (steps)")
update(splot,main="Comparison of Average Number of Steps During 5-minute\nTime Intervals In a 24-hour Day\nfor Weekend Days Versus Weekdays")
```
Weekday steps show a large peak around 8:40AM followed by 4 smaller peaks around lunch time, afternoon break time, and supper time. Step data appears more uniform throughout weekend days and have smaller peaks.





